---
title: "Assignment 5 Solutions"
date: "Thursday March 17, 2016"
author: Dexter barrows
output:
  html_document:
    theme: cosmo
    highlight: tango
---

```{r, echo = FALSE, warning = FALSE}
library("knitr")
opts_chunk$set(tidy = FALSE, engine = 'python',engine.path = 'python3')
```



***

## Question 1

Write a function `array_slicemean(a)` that takes a given 3-D array (such as `aa=np.arange(27).reshape((3,3,3))`) and returns a vector containing the sum of each slice of the *first* dimension. For example, the result of applying `array_slicemean` to the example given here is `np.array([ 36, 117, 198])`.

**Loop way**

```{r}
import numpy as np

def array_slicemean(a):
	res = []
	for i in range(a.shape[0]):
		slicesum = 0
		for j in range(a.shape[1]):
			for k in range(a.shape[2]):
				slicesum += a[i,j,k]
		res.append(slicesum)
	return( np.array(res) )

a = np.arange(27).reshape((3,3,3))
print( array_slicemean(a) )
```

**Numpy way**

```{r}
import numpy as np

def array_slicemean(a):
	return( np.sum(a, axis = (1,2)) )

a = np.arange(27).reshape((3,3,3))
print( array_slicemean(a) )	
```


***

## Question 2

Write a function `array_poly(a,x)` that takes a 2-dimensional array `a` and a scalar (floating-point) `x` and computes a vector of polynomials using each row as a separate set of coefficients. In other words, if `a` has three columns, then the `i`^{th} element of the returned array would be `a[i,0]+a[i,1]*x+a[i,2]*x**2`.

**Loop way**

```{r}
import numpy as np

def array_poly(a, x):
	vec = []
	for i in range(a.shape[1]):
		vec.append(x**i)
	res = []
	for i in range(a.shape[1]):
		polysum = 0
		for j in range(a.shape[0]):
			polysum += a[i,j]*vec[j]
		res.append(polysum)
	return( np.array(res) )

a = np.reshape(np.arange(9),(3,3))
print( array_poly(a,2) )
```

**Numpy way**

```{r}
import numpy as np

def array_poly(a, x):
	vec = x**np.arange(a.shape[1])
	return( np.dot(a, vec) )

a = np.reshape(np.arange(9),(3,3))
print( array_poly(a,2) )
```


***


## Question 3

Write a function, `check_inverse(a,b,tol=1e-8)` that takes two 2-dimensional square arrays (matrices) and returns a boolean value that reflects whether `b` is the matrix inverse of `a`, i.e. that `numpy.dot(a,b)` is equal, within tolerance `tol`, to an identity matrix of the same dimension (use `np.eye()`). Make sure to:

- check that the arrays are both square and have the same dimensions, and raise a `ValueError` if not
- use an appropriate test for floating-point near-equality; that is, return `True` if, for every $i$ and $j$, $\textrm{abs}((AB)_{ij}-I_{ij})<\textrm{tol}$, and `False` otherwise.

**Loop way (with a little Numpy)**

```{r}
import numpy as np

def check_inverse(a, b, tol = 1e-8):
	if a.shape[0] != a.shape[1] or b.shape[0] != b.shape[1] or a.shape[0] != b.shape[0]:
		raise ValueError
	mat = abs( np.dot(a, b) - np.eye(a.shape[0]) )
	for i in range(mat.shape[0]):
		for j in range(mat.shape[1]):
			if mat[i,j] >= tol:
				return(False)
	return(True)

np.random.seed(1)
a = np.random.random(9).reshape((3,3))
b = np.linalg.inv(a)
print( check_inverse(a,b) ) 	# Should be True
print( check_inverse(a,b,0) )	# Should be False
print( check_inverse(a,b+2) )	# Should be False
```

**Pure Numpy**

```{r}
import numpy as np

def check_inverse(a, b, tol = 1e-8):
	if a.shape[0] != a.shape[1] or b.shape[0] != b.shape[1] or a.shape[0] != b.shape[0]:
		raise ValueError
	return( np.all( abs(np.dot(a,b)-np.eye(a.shape[0])) < tol ) )

np.random.seed(1)
a = np.random.random(9).reshape((3,3))
b = np.linalg.inv(a)
print( check_inverse(a,b) ) 	# Should be True
print( check_inverse(a,b,0) )	# Should be False
print( check_inverse(a,b+2) )	# Should be False
```


***

## Question 4

Define a function `maxrows(a,m)` that returns an array comprising all the *rows* in the array whose mean is greater than `m`. If we have the usual `a = np.arange(9).reshape((3,3))` then the result of `maxval(a,2.1)` should be `np.array([[3, 4, 5], [6, 7, 8]])`.

**Loop way**

```{r}
import numpy as np

def maxrows(a, m):
	res = []
	for i in range(a.shape[0]):
		rowsum = 0
		for j in range(a.shape[1]):
			rowsum += a[i,j]
		rowmean = rowsum / a.shape[1]
		if rowmean > m:
			res.append(a[i])
	return( np.array(res) )

a = np.arange(9).reshape((3,3))
print( maxrows(a,2.1) )
```

**Numpy way**

```{r}
import numpy as np

def maxrows(a,m):
    v = np.mean(a, axis=1) > m
    return(a[v])

a = np.arange(9).reshape((3,3))
print( maxrows(a,2.1) )
```