### Benford's Law

Benford's law is ...

Read it about it on [Wikipedia](http://en.wikipedia.org/wiki/Benford%27s_law) or
[MathWorld](http://mathworld.wolfram.com/BenfordsLaw.html)

Write a Python function `benford_count` that takes the name of a file (that will contain one number per line), opens the file, and tabulates the occurrences of last digits on each line.

**NOTE** that if you follow the instructions below to read and process the file, the last character of each line is a *newline* character; if you want the last *digit* you need to get the second-to-last character of the string.

```{r eval=FALSE}
ff = open("benford.txt")
for line in ff:
    ## code
```

For example, if the text file contained only three lines
```
1,234
123
12
```
then the correct result would be `[0,0,1,1,1,0,0,0,0,0]` (i.e. the last digits per line in this file comprise one 2, one 3, and one 4; there are zero of every other digit).

Keep in mind that (1) what you read out of the file will be a *string* and (2) you can turn a digit into an integer using `int()`, e.g.

```{r}
x = "7"
print(int(x))
```
