---
title: "math 1MP assignment 4"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

### Logistics

Homework #4 is due at midnight (i.e. 11:59:59 PM) on Weds February 11.

You **must** submit your homework to the Dropbox on Avenue to Learn as a **single** text (`.py`) file containing two function definitions (`int_to_roman` and `logist_map`, as specified below).  The file must be named `[MacID]_hw4.py`, e.g. in my case it would be called `bolkerb_hw4.py`.  You will automatically lose at least one mark if you don't follow these instructions ...

The challenge questions below are completely optional and *not* good for extra credit.

### Converting integers to Roman numerals

We've spent a lot of time in class talking about how to convert a Roman numeral to an integer.  Now you should write a function to convert integers to Roman numerals.  The rules are:

* 1000 $\to$ M
*  900 $\to$ CM
*  500 $\to$ D
*  400 $\to$ CD
*  100 $\to$ C
*   90 $\to$ XL
*   50 $\to$ L
*   40 $\to$ XL
*   10 $\to$ X
*    9 $\to$ IX
*    5 $\to$ V
*    4 $\to$ IV
*    1 $\to$ I

The symbols should be arranged from largest to smallest (with the exception of 'subtraction rule' cases), e.g. 1909 $\to$ "MCMIX".

The key to this problem is using the operators `//` (integer division) and `%` (modulo, i.e. the remainder of integer division).  For example,
```{r intdiv}
print(1909 // 1000)
print(1909 % 1000)
```
If you use these operators, you can process the value/symbol pairs above one at a time and not worry about back-tracking or figuring out whether the subtraction rule is being used.

The operators for adding (concatenating) strings and multiplying them by integers will be useful too, e.g.
```{r str_add}
print("V"+"I"*3)
```

Write a Python function `int_to_roman` that takes a single argument `n` and converts it to a character variable holding the corresponding Roman numeral.  You can assume that the input will be an integer between 1 and  3000, inclusive.

You can use brute force if you like but it will probably be easier to set up a pair of lists (`symbols` and `values`) that contain the symbols and their corresponding value. 

**Challenge**: do this problem with a smaller set of symbol/value tables that doesn't use the subtraction-rule pairs (i.e. the symbols are only `["I","V","X","L","C","D","M"]`).

### The discrete logistic

The *discrete logistic map* is the name for the iterative mathematical equation
$$
x_{t+1} = r x_t (1-x_t)
$$
This equation has very interesting properties which you can [read about on Wikipedia](http://en.wikipedia.org/wiki/Logistic_map) if you like.

Write a Python function called `logist_map` that takes the parameters `x0` (the starting value), `r` (the growth rate parameter), `t_trans` (the "transient time", i.e. the number of time steps to throw away), and `t_save` (the number of time steps to save).  The function should compute the discrete logistic map starting at `x0`.  For example,  if `r`=0.5, `x0`=0.5, `t_trans`=2, and `t_save`=3, then we would compute the sequence by hand as follows:

```{r log1}
x0 = 0.5; r=0.5
## take 2 steps
x1 = r*x0*(1-x0)
x2 = r*x1*(1-x1)
## now take 3 more steps
x3 = r*x2*(1-x2)
x4 = r*x3*(1-x3)
x5 = r*x4*(1-x4)
## print the result
print([x3,x4,x5])
```

If this were a function you would use:

```{r log2}
def logist_map(x0,r):
    ## take 2 steps
    x1 = r*x0*(1-x0)
    x2 = r*x1*(1-x1)
    ## now take 3 more steps
    x3 = r*x2*(1-x2)
    x4 = r*x3*(1-x3)
    x5 = r*x4*(1-x4)
    ## return the result
    return([x3,x4,x5])

print(logist_map(x0=0.5,r=0.5))
```

Of course, since you won't know `t_trans` or `t_save` in advance you will need to use `for` or `while` loops as appropriate.

Your program should check and raise a `ValueError` if:

* `r` is <0 or >4 (these bounds make sure that the `x` value will always stay bounded between 0 and 1, if it starts there)
* `x0` is <0 or >1

You can assume that `r` and `x0` are numeric and that `t_trans` and `t_save` are non-negative integers.

**Challenge question**: think about how you would take a list of results from the logistic map and try to estimate whether `x` is changing with some periodicity.
