---
title: "math 1MP assignment 5"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

### Logistics

Homework #5 is due at midnight (i.e. 11:59:59 PM) on Weds March 4.

You **must** submit your homework to the Dropbox on Avenue to Learn as a **single** text (`.py`) file containing five function definitions [`setup_dict()`; `update_dict()`; `count_file()`; `odd_matrix()`; `flipadd_matrix()` as specified below]. Make sure your file also contains appropriate `import` specifications for the modules you use. The file must be named `[MacID]_hw5.py`, e.g. in my case it would be called `bolkerb_hw5.py`.  You will automatically lose at least one mark if you don't follow these instructions ...


### Problem 1: Dictionary use

Write functions to compute letter frequencies from a document.

* Write a function `setup_dict()` that returns a dictionary whose keys are all of the lowercase letters, each with value zero
    * you may use `import string; letters=list(string.ascii_lowercase)` to get a list of lowercase letters
* Write a function `update_dict(d,line)` that takes a line, converts it to lowercase (using `str.lower()`) and, *ignoring values that are not in the dictionary*, modifies the dictionary to add the counts of the letters in the line 
    * *hint*: you can use `in` to check whether a key is present in a dictionary
	* your function should return the updated value of the dictionary.
* Write a function called `count_file(fn)` that takes a file name (e.g., `"dat.txt"`) and returns a dictionary of letter counts from it.  
    * It should work by calling `setup_dict()` and `update_dict()` appropriately.
	* Use `f = open(fn)` to open the file; you can then run `for line in f:` to extract successive lines into the variable `line`.

### `numpy` problems

#### 2.

* create a function `odd_matrix(n)` that returns an n-by-n array containing the squares of the first $n^2$ odd numbers *in column-major order*, i.e. `odd_matrix(2)` is
```
1  25
9  49
```
while `odd_matrix(3)` is
```
1  49   169
9  81   225
25 121  289
```
* *hint*: `numpy.reshape(a,(r,c))` turns an array into an $r \times c$ matrix (i.e., a 2D array with `r` rows and `c` columns), but it does it in row-major order ...

#### 3.
* create a function `flipadd_matrix(m)` that, given a 2D array `m` with `c` columns and `2*c` rows:
    * creates a new matrix by concatenating `m` with itself column-wise; 
	* creates a new matrix by transposing `m` and concatenating the result with itself row-wise;
    * adds the result to return an array with `2*c` rows and `2*c` columns.
	* For example, if
```{r a1,eval=FALSE}
m = np.array([[0],
              [1]])
```
We first concatenate it with itself column-wise to get:
```{r a1b,eval=FALSE}
array([[0, 0],
       [1, 1]])
```
Starting with `m` again, we transpose (to get `array([[0,1]])`, i.e. a 1 $\times$ 2 array) and then concatenate row-wise to get:
```{r a1c,eval=FALSE}
array([[0, 1],
       [0, 1]])
```
When we add the two arrays elementwise above, the resulting matrix
(which your `flipadd_matrix(m)` should return) is:
```{r a2,eval=FALSE}
array([[0, 1],
       [1, 2]])
```
Example #2: if
```{r a3,eval=FALSE}
m = np.array([[0, 1],
              [2, 3],
              [4, 5],
              [6, 7]])
```
then the result would be
```{r a4,eval=FALSE}
array([[ 0,  3,  4,  7],
       [ 3,  6,  7, 10],
       [ 4,  7,  8, 11],
       [ 7, 10, 11, 14]])
```
* *hint*: `numpy.concatenate()` takes a *tuple* of arrays as its first argument, and an optional `axis` argument to specify the direction of concatenation (0=row-wise, 1=column-wise).
