---
title: "math 1MP term projects"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

These projects are **individual** projects; if you discuss anything about them with other students in the class, you should follow the standard rules (don't write anything down while you are engaged with the other students).  If you get anything from the web, give references.

### Yahtzee

Yahtzee is a dice game.  One small subset of the rules specifies that you have three rolls of five (six-sided) dice in which to try to get as many of a kind as possible.  You only have to re-roll as many dice as you want, which means that if you have at least two of a kind you can maximize your chances of getting "Yahtzee" (five of a kind) by not re-rolling those dice.  Write the following simulation functions:

```{r
def single_roll(n,fixed=[],sides=6):
    """perform a single roll of n dice (default 6-sided)
       combine the results with the fixed values
       return a list containing the result
    """

def new_roll(vals,sides=6):
    """given a list of die values, pick which to keep
       fixed and which to re-roll
       and re-roll them"""

def sim_yahtzee(nrolls,sides=6):
    """simulate a specified number of dice rolls.
       Return a boolean True/False value depending on whether
       you achieve all dice the same
    """

def est_yahtzee_prob(nsims,nrolls=3,dice=5,sides=6):
    """estimate the probabilities of Yahtzee for a specified number
       of rolls, dice, and sides"""
```

Put these functions together to compute the probability of Yahtzee
(use 100,000 simulations each) for 

* the standard case (3 rolls of 5 6-sided dice)
* 2 to 6 rolls (with the standard 5 6-sided dice)
* 5, 6, or 7-sided dice (in 3 rolls of 5 dice).  
* Approximately how many rolls do you need with 5 7-sided dice to get close to the probability in the base case?

### n-grams

This is a variant of [this project](http://www.cse.msu.edu/~cse231/PracticeOfComputingUsingPython/06_Dictionaries/MarkovTextBuilder/proj07.pdf).

An "n-gram" is a set of `n` consecutive words from a document or string; for example, the phrase "now is the time for all good people" contains six 3-grams: 
```{r echo=FALSE}
test = "now is the time for all good people"
from markov import *
a = add_ngrams(test.split(),n=3)
print([i + tuple(a[i]) for i in a])
```
(note that these are not in order; I stored them in a dictionary ...)

The first part of this project is to tabulate the n-grams in a document.

The structure of the set of n-grams should be a *dictionary* whose keys are tuples of `n-1` words, and whose contents are lists of the next strings found in the document.  For example, if we have the string "today is my birthday today is also his birthday" and we want to tabulate the 2-grams, we would get this:
```{r echo=FALSE}
test = "today is my birthday today is also his birthday"
from markov import *
print(add_ngrams(test.split(),n=2))
```

```{r}
def get_file(fn):
    """read in an entire file as a single string"""
    f = open(fn)
    r = f.read()
    f.close()
    return(r)

def clean_string(str):
    """remove all punctuation and newlines ('\n') from a string,
       and make it lowercase"""
     ## hint: str.remove(), str.lower()

def make_ngram_dict(strList,n=2):
    """make an n-gram dictionary from a list of strings"""
    ### hint: clean_string(), str.split()
    ### check whether tuples of (n-1) words are present in the dictionary
    ###    if not, add an entry with the next word
    ###    if so, add the word to the list of words following that tuple
```

The second part of the project is to make a *Markov generator*. Given 
a starting set of `n-1` words, pick the next word at random from the
appropriate dictionary entry (`numpy.random.choice(x)` will pick a
random element from the list `x`) and add it to the list of words.
If the last word in the text you used to set up the dictionary is unique,
then the chain will stop at that point: return `None`.

```{r eval=FALSE}
def next_word(cur_list,n=2,ngram_dict):
    """use an ngram dictionary to generate the next word in a list"""
    ##

def create_list(nwords,start_list,n=2,ngram_dict):
    """use an ngram dictionary a list of words"""
    ##


```
