---
title: "Tutorial 8(ish): Numpy"
date: "Tuesday January 19, 2016"
author: Dexter barrows
output:
  html_document:
    theme: cosmo
    highlight: tango
---

```{r, echo = FALSE, warning = FALSE}
library("knitr")
opts_chunk$set(tidy = FALSE, engine = 'python',engine.path = 'python3')
```

[Documentation reference](http://docs.scipy.org/doc/numpy/reference/).



***

## Numerical Types Revisited

Recall that numbers in Python are either of type `int` or `float`. To expand on that a little:

#### IntN

* Uses N bits (N/8 bytes) to store each number using [binary representation](https://en.wikipedia.org/wiki/Binary_number#Counting_in_binary)
* Stores whole numbers in the range $\left[ -2^{N-1},2^{N-1}-1 \right]$ (one bit used for sign)
* On most machines $N = 64$, older machines might use $N = 32$

#### FloatN

* Again uses N bits to store each number, but using an [exponent and mantissa/significand](https://en.wikipedia.org/wiki/Floating_point) (one bit still used for sign)
* Entirely possible that a number will be either too big or small to use in arithmetic, called 'overflow' and 'underflow'
* Smallest usable number can be determined using Numpy

```{r}
import numpy as np
print( np.finfo("float64").eps )
```

So if we try something smaller, say $2^{-17}$

```{r}
print( 1.0 + 2e-17 )
```

Underflow!

What about overflow? We can get the smallest power of 2 that yields overflow.

```{r}
import numpy as np
mexp = np.finfo("float64").maxexp
print( 1.0 + 2.0**(mexp-1) )
```

But if we do

```{r}
import numpy as np
mexp = np.finfo("float64").maxexp
try:
	print( 1.0 + 2.0**mexp )
except Exception as e:
	print(e)
```

Overflow!


***


## Arrays and Matrices

#### 1D

We have been using `lists` to represent arrays of numbers, but it's slower and really less robust for actual *math*.

Say we have a `list` we want to convert to an array so we can perform more advanced arithmetic:

```{r}
import numpy as np
L = [1,2,3,4,5]
A = np.array(L)
print(L)
print(A)
```

Now we can perform element-wise operations like adding or multiplying

```{r}
import numpy as np
A = np.array([1,2,3,4,5])
print( A+A )
print( A*A )
```

We can also create arrays directly using `numpy.zeros()`, `numpy.ones()`, `numpy.arange()`, etc.

```{r}
import numpy as np
A0 = np.zeros(5)
A1 = np.ones(5)
AR = np.arange(5)
print(A0)
print(A1)
print(AR)
```

**Indexing** and **slicing** work the same as with `lists` / `strings` / `tuples`.

#### 2D

We can also have 2D arrays (matrices). We can create them using a nested `list` / `tuple`. 

```{r}
import numpy as np
M = np.array( [[1,2],[3,4]] )
print(M)
```

As before we can perform element-wise operations.

```{r}
import numpy as np
M = np.array( [[1,2],[3,4]] )
print( M+M )
print( M*M )
```

Indexing / slicing now uses `[i,j]` for row/column instead of `[i][j]`.

```{r}
import numpy as np
M = np.array( [[1,2],[3,4]] )
print( M[0,] )
print( M[1,0] )
print( M[-1,-1] )
```

Note that leaving the column index blank gives the entire row. This does **not** work the other way around -- `M[,0]` with the matrix above would give a Syntax Error.

We can also create matrices directly using `numpy.eye()`, or `numpy.zeros()`/`numpy.ones()` as before -- just remember to specify the size of *each* dimension.

```{r}
import numpy as np
M0 = np.zeros([2,2])
M1 = np.ones([2,3])
ME = np.eye(5)
print(M0, end = "\n\n")
print(M1, end = "\n\n")
print(ME, end = "\n\n")
```

***

## Exercises

[http://bbolker.github.io/math1mp/notes/numpy_ex.html](http://bbolker.github.io/math1mp/notes/numpy_ex.html)