---
title: "Assignment 4 Submission Notes"
date: "Tuesday March 1, 2016"
author: Dexter Barrows
output:
  html_document:
    theme: cosmo
    highlight: tango
---

```{r, echo = FALSE, warning = FALSE}
library("knitr")
opts_chunk$set(tidy = FALSE, engine = 'python',engine.path = 'python3')
```

***

## Filename

As stated in the assignment instructions, your filename **must** be in the format `yourmacid_hw4.py`.

Incorrect filename formats that I have seen:

* `yourmacid.py` - missing `_hw4`
* `yourmacid_HW4` - capitals
* `yourmacid_hm4.py` - should be `hw4` not `hm4`
* `yourmacid_hw4(py).txt` - wrong extension, brackets

You **will lose marks** if your filename is not in the correct format.

***

## File Contents

Your submission **must contain only functions** and **nothing else**.

For example, if my file was named `barrowdd_hw4.py` and contained:

```{r, eval = FALSE}
def add_bigram(d, key, w):
	"""Docstring for add_bigram"""
	...
	...

def create_bigram_dict(fn):
	"""Docstring for create_bigram_dict"""
	...
	...
```

then this would be correct. If you want, you can test your code by adding a check to determine if the program is the main one or if it has been imported as a module (which is how I check them).

This looks like:

```{r, eval = FALSE}
def add_bigram(d, key, w):
	"""Docstring for add_bigram"""
	...
	...

def create_bigram_dict(fn):
	"""Docstring for create_bigram_dict"""
	...
	...

if __name__ == "__main__":
	fn = "darwin_nopunct.txt"
	dd = create_bigram_dict(fn)
	print( dd["so"] )
```

The idea is that anything after the `if __name__ = ...` check will only be executed if you are running it, but not if I am running it.

***

## Improper Collaboration

There were many submissions for Assignment 3 that were obvious examples of improper collaboration (I have a list). We are giving you the benefit of the doubt **once**. Further suspicious submissions will result in a mark of 0.