---
title: "Extra Python Practice"
date: "Tuesday February 2, 2016"
author: Dexter barrows
output:
  html_document:
    theme: cosmo
    highlight: tango
---

```{r, echo = FALSE, warning = FALSE}
library("knitr")
opts_chunk$set(tidy = FALSE, engine = 'python',engine.path = 'python3')
```

[Python documentation reference](https://docs.python.org/3/tutorial/controlflow.html).

***

## Quick Functions

When you do something like this

```{r}
numlist = [1,2,3]
print( sum(numlist) )
```

you are calling two functions, `print()` and `sum()` that are built into Python.

You can define your own functions using this structure:

```{r, eval = FALSE}
def function_name(arg1, arg2, ...):
	body_1
	body_2
	...
	return value
```

For example, an alternative to the built in `sum()` function:

```{r}
def dex_sum(alist):
	total = 0
	for a in alist:
		total += a
	return total

numlist = [1,2,3]
print( sum(numlist) )
print( dex_sum(numlist) )
```

Note that for **all** functions:

- `def` is keyword to tell Python we are defining a function
- `return` is used to return a value or object

And for this example in particular:


- `dex_sum` is the name of the function you are defining
- `alist` is the argument passed into the function



***

## CodingBat

[CodingBat](http://codingbat.com/python) is a site where you can practice Python, similar to [CodeLab](http://www.turingscraft.com/go.html).

All CodingBat exercises ask you to fill in the body of a function (see above).

Please try to do (some of these are a bit tricky):

- [Logic 1](http://codingbat.com/python/Logic-1)
- [List 2](http://codingbat.com/python/List-2)
- [String 2](http://codingbat.com/python/String-2)

If you get bored and/or want more of a challenge:

- [Logic 2](http://codingbat.com/python/Logic-2)