---
title: "Factoring, Euclidean algorithm and rhythms"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine="python",engine.path='python3')
```

## The Euclidean algorithm

### definition and history

* oldest known algorithm (Euclid, c. 300 BCE) ("algorithm" from al-Khwārizmī, c. 780-850 CE)
* find **greatest common divisor** of two integers: gcd(`a`,`b`) is the largest integer `c` such that `c` divides both `a` and `b` evenly.
* ex.: gcd(36,42)=6
* brute force: enumerate factors of `a` and `b`, multiply common factors ($36=2^2 \times 3^2$, $42 = 2 \times 3  \times 7$)
```{r}
def find_factors(a):
    """find integer factors (other than 1 and a) by brute force"""
    factors = []
    i = 2
    while i<=a:
        ## print(i,a%i,a//i,factors)
        if a%i==0:
            factors += [i]
            a = a//i
        else:
            i += 1
    return(factors)

f1 = find_factors(36)
f2 = find_factors(42)
def common_factors(f1,f2):
    for i in f1:
        if not (i in f2):
            f1.remove(i)
    for i in f2:
        if not (i in f1):
            f2.remove(i)
    if (len(f1)<len(f2)):
        return(f1)
    else:
        return(f2)
print(common_factors(f1,f2))
```

* if the gcd is 1, the numbers are **co-prime** (but not necessarily both prime!)

### Subtractive form

* subtract the smaller number from the larger; take the smaller and the remainder and repeat, until you get to zero
* $\{42,36\} \to \{6,0\}$
* $\{407,361\} \to \{361,46\} \to \{315,46\} \to ... \to \{85,46\} \to \{46,39\} \to \{39,7\} \to \{32,7} ... \to \{7,4\} \to \{4,3\} \to \{3,1\}$ ... co-prime!
* intuition
* code:
* looping:
```{r euc_sub_loop}
def euc_alg1(a,b):
    '''subtractive Euclid algorithm by looping'''
    while b>0:
        lg = max(a,b)
        small = min(a,b)
        b = lg-small
        a = small
        ## print(a,b)
    return(a)
```    
* recursive:
```{r euc_sub_recurse}
def euc_alg2(a,b):
    '''subtractive Euclid algorithm by recursion'''
    if (b==a):
        return(a)
    return(euc_alg2(min(a,b),abs(a-b)))
```

### division form
* intuition: there's no need to subtract repeatedly when we could do the same thing in one step by dividing and taking the remainder.  
* example:
    * $\{42,36\} \to \{6,0\}$ (one step)
    * $\{407,361\} \to \{361,46\} \to  \{46,39\} \to \{39,7\} \to \{7,4\} \to \{4,3\} \to \{3,1\}$ ... co-prime!
* looping
```{r euc_div_while}
def euc_alg3(a,b):
    '''modular Euclid algorithm by looping'''
    while b>0:
       r = a % b
       a = b
       b = r
       print(a,b)
    return(a)
```
* recursive
```{r euc_div_recurse}
def euc_alg4(a,b):
    '''modular Euclid algorithm by recursion'''
    if (b==a):
        return(a)
    m = min(a,b)
    return(euc_alg2(m,max(a,b) % m))
```

## Computational complexity and timing

* count operations
* want to know how complexity **scales** with problem size
* e.g. $O(n)$, $O(\log n)$, $O(\exp(n))$ (uh-oh!)
* polynomial-time vs. non-polynomial
* best-case vs average vs worst-case

## Bjorklund algorithm

### Worked example

## References

* old paper on EA applications (in math): http://www.jstor.org/stable/3029367
* http://www.hisschemoller.com/blog/2011/euclidean-rhythms/
* http://cgm.cs.mcgill.ca/~godfried/publications/banff-extended.pdf
* Python implementations: (watch out for Python 2 constructions!  need to replace `/` by `//` for integer division; replace `xrange()` with `range()` ...
    * http://www.atonalmicroshores.com/2014/03/bjorklund-py/
    
    
