---
title: "the Collatz conjecture"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Ben Bolker
---


```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

From [Wikipedia](https://en.wikipedia.org/wiki/Collatz_conjecture):

> Take any natural number n. If n is even, divide it by 2 to get n / 2. If n is odd, multiply it by 3 and add 1 to obtain 3n + 1. Repeat the process (which has been called "Half Or Triple Plus One", or HOTPO[6]) indefinitely. The conjecture is that no matter what number you start with, you will always eventually reach 1. 

```{r colfun}
def collatz(n,itmax=10000):
    it = 0
    while not n==1 and it<itmax:
       it += 1
       if n % 2 == 0:
           n = n//2
       else:
           n = 3*n+1
       print(n)
    print(str(it)+" steps")

collatz(4)
collatz(10)
```

Some questions to think about:

- when is the sequence of minimal length?
- how does the length grow with the size of the initial $n$?

But [be careful](https://xkcd.com/710/)!
