---
title: "Slicing with a negative stride"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
## sigh, still doesn't work
## library("runr")
## py <- proc_python()
## py$start()
## knit_engines$set(python = function(options) {
##      knitr:::wrap(py$exec(options$code), options)
## })
```

As mentioned in class, using a slice expression with three colons on a string (e.g. `s[a:b:n]` uses `a` as the starting point; `b` as the stopping point; and `n` as the *stride*, or the step. If you leave out one of `a`, `b`, or `n`, Python fills in a sensible default: for forward stepping (positive `n`), this is `a=0` and `b=len(s)`.

```{r s}
s = "Monty Python"
print(s[::])   ## leave out a, b, n: means 'take the whole string'
               ## (start=0, stop=len(s) (take up to position len(s)-1), stride=1)
print(s[0:len(s):1]) ## the same, but explicit
print(s[2::2])       ## start at position 2, every other character thereafter
print(s[:10:2])      ## start at the beginning (0), every other character up to #9
```

If we use a *negative* stride (i.e., go backwards), then starting at position 0 and stopping at position `len(s)-1` will get us no characters at all: 

```{r s2}
s = "Monty Python"
print(s[0:len(s):-1])
```

```{r s3}
s = "Monty Python"
print(s[len(s):0:-1])
```
