---
title: "digits of powers of 2"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Ben Bolker
---

Random thought #1, for an interesting math programming problem. *What is the distribution of digits of powers of 2?*

Are they uniform, or do they have some interesting pattern?  At the beginning (2,4,8,16) the sequence is odd-free; the first "3" occurs at 2^5=32, the first "5" at 2^8=256, the first "7" at 2^15 (but the first "9" at 2**12).  Does the rarity of odd digits continue?

In its simplest form, this is a very easy program to write: you just need to loop over powers of 2 and use string processing to accumulate digits.

The patterns are probably more interesting if we keep track of them by digit. For example, it's trivial to figure out the least-significant digit patterns, and not too hard to figure out the most-significant digit patterns:

```{r engine="python"}
res = []  ## empty list
for x in range(20):
    res += format(2**x)[-1]
print(res)
```

You can also do this slightly more compactly with a more-advanced Python feature called a [list comprehensions](https://docs.python.org/2/tutorial/datastructures.html):

```{r engine="python",results="hide"}
[format(2**x)[-1] for x in range(20)]
```

The most significant digits also have a pattern, although it's a little bit less obvious:
```{r engine="python"}
print([format(2**x)[0] for x in range(20)])
```

Next things to do/extensions:

* store the results as a list averaged by number of digits, so that we can go through and pick out the first/last/second-to-last etc. digit.
* draw pictures
* think about *computational complexity* of this problem, and do some timing to test the conjecture that it's $O(n \log n)$
