---
title: "square roots of negative numbers"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Ben Bolker
---

```{r opts,echo=FALSE,message=FALSE}
## source("python3_engine.R")
knitr::opts_chunk$set(engine="python",engine.path="python3")
```

I was wondering whether there was an easier way or a shortcut for writing `complex(a,b)` to get a complex number of the form $a+bi$ (or $a+bj$, in keeping with Python's use of `j` for $\sqrt{-1}$.  It took me a little while, but the way that I tried to get there may be interesting.

First, I tried to take the square root of -1:
```{r sqrtneg1}
print(-1**0.5)
```
This didn't work due to a logical but surprising glitch: the *operator precedence* (i.e., the order in which operators get executed) is higher for the power operator `**` than for the **unary minus** operator `-`.  In other words, Python interprets `-1**0.5` as `-(1**0.5)`, not `(-1)**0.5`.

Then I guessed that there might be a square-root function called `sqrt()`: this is common in a lot of computer languages

```{r sqrtfun,error=TRUE}
sqrt(-1)
```

Using a little bit of additional information (I knew/guessed that there is an extra library, or *module*, in Python called `math`, and that the following is the way to access it):

```{r mathsqrtfun1,error=TRUE}
import math     ## load the math module
math.sqrt(-1)
```

OK, that doesn't work. Apparently `math.sqrt` defines a function whose domain is restricted to non-negative values (that's what `math domain error` means). What if I try reading the help file?
```{r help,eval=FALSE}
help("complex")
```
That showed me that there is indeed a `pow` function for complex numbers (admittedly this is a little hard to read!), but that's not really the point -- I want to take `pow()` of a *negative* floating point number with a *fractional* power.

Just out of curiosity, what happens if I try `math.sqrt()` on a number that's already defined as complex?

```{r mathsqrtfuncomp,error=TRUE}
math.sqrt(complex(-1,0))
```

Oh well, it looks like the `math` module is really only designed for floating-point operations, not operations on complex numbers ...

Coming back to what (in hindsight) I should have done in the first place, I tried (1) the `pow()` function (which takes care of the order-of-operations problem; `pow()` is definitely applied *after* the unary minus sign and (2) making sure I put parentheses around the `(-1)` when using the `**` operator.

```{r pow}
print(pow(-1,0.5))
print((-1)**0.5)
```

These both work (**but only in Python 3**: Python 2 gives "negative error cannot be raised to a fractional power"). They don't give me *exactly* what I wanted, though. Python tells me the answer is $\epsilon+i$, where $\epsilon$ is a tiny number (not quite zero).  (This kind of problem is an inevitable consequence of trying to approximate real numbers via finite binary approximations; we will talk about it more in a few weeks.) The R language has a `zapsmall` function that sets small values to zero; the `numpy` module for Python has a `real_if_close()` function, but that's not quite what we want (we want to set the *real* part to zero).

If I really wanted to, I could set `i` (or `j`) equal to `complex(0,1)`, or to `pow(-1,0.5)`, and then define complex variables via `real_part+i*imag_part` (e.g. `2.72+0.65*i`), but it would only be worth it if I were working with complex variables a lot.

And **now**, just after I've finished with this whole thing, I've realized that there was a shortcut all along:

```{r final}
print(2.72+0.65j)
```
works just fine.

This example illustrates (1) the potential subtlety of order of operations; (2) there's more than one way to do it; (3) how to experiment/look for help; (4) read/interpret error messages.
