---
title: "generating list copies"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Ben Bolker
---

```{r opts,echo=FALSE,message=FALSE}
## source("python3_engine.R")
knitr::opts_chunk$set(engine="python",engine.path="python3")
```

How would you generate a list comprising `n` `[1]` values and `k-n` `[0]` values as compactly as possible (1) *without* shallow copying within the results and (2) *without* using list comprehensions?  (The hard part of this question is that the elements of the list are lists; if they were immutable types such as numbers the multiplication solution would work, because copying the template would make deep copies.)

  Here are three solutions:

```{r defs}
def list_mult(n,k):
    return([[1]]*n+[[0]]*(k-n))
def list_comp(n,k):   
    return([[1] for i in range(n)] + [[0] for i in range(k-n)])
def list_for(n,k):
   res = []
   for i in range(n):
      i+=[1]
   for i in range(k-n):
      i+=[0]
   return(res)
```

* `list_mult` makes shallow copies (so we can't modify elements independently);
* `list_comp` uses list comprehensions;
* `list_for`  feels clunky.

Maybe this is an unfair set of constraints; in other words, the answer might be "list comprehensions are the right way to do this, get over it, *or* write a utility module that encapsulates `list_for` and import it so you don't have to see the complexity"

Another possible solution would be to use lists of tuples instead of lists of lists, since tuples are immutable.

```{r}
def list_tuple(n,k):
    return([(1,)]*n+[(0,)]*(k-n) )
```

The only downside I see of this is that the syntax of `[(1,)]` is a little bit inscrutable ("make a list containing a single element which is a single-element tuple containing 1") -- starts to approach [code golf](http://en.wikipedia.org/wiki/Code_golf).  Maybe this is just because single-element tuples look weird (they have to, because `(1)` would be indistinguishable from a parenthesized number).
