---
title: "assignment"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

I was surprised by the picture on p. 20 of Gries *et al.*, which shows a variable (`difference`) that was previously set equal first to 26 and to 5 pointing at a new memory location containing 5, while the previous memory location is still there. I would have expected (based on C reflexes) that the same memory location would be overwritten by the new value, but this is not true: instead, Python allocates (semi-persistent) memory for the literal values it sees.

```{r idvars}
x = 1
print(id(x))
```

```{r idvars2}
print(id(1))
```

```{r idvars3}
x = 2
print(id(x))
print(id(1))
```

Voilà, `x` is pointing at a new memory location, `1` is still sitting at the same old memory location.

This is (perhaps better) explained [here](http://foobarnbaz.com/2012/07/08/understanding-python-variables/), with pretty pictures:

> Other languages have 'variables'. Python has 'names'.
