---
title: "regular expressions"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

Large topic -- somewhat more advanced than 'basic programming', but worth a digression.  ([reference](https://docs.python.org/3/howto/regex.html))

What if we are looking for some number, but we don't know what number?
```{r re}
import re
bool(re.search('[0-9]', 'Plan 9'))
```

----------------------------------------------------------
   Pattern   Description                                           
-----------  ---------------------------------------------
 `^`         Beginning of line
 
 `$`	     End of line    
 
 `.`         Any single character except newline
 
 `[...]`	 Any single character in brackets         
 
 `[^...]`	 Any single character **not** in brackets 
 
 `re*`	     0 or more occurrences of preceding expression
 
 `re+`	     1 or more occurrence of preceding expression
 
 `re?`	     0 or 1 occurrence of preceding expression  
 
 `re1|re2`   match `re1` or `re2`
 
 `()`        grouping
 
----------------------------------------------------------

* How would you test whether a string contains a numeric value at the end (e.g. "Plan 99")?
* What if the string might contain a comma (e.g. "Plan 99,478")?
* What if you're looking for the abbreviations of rooms in Hamilton Hall (my office is HH314)?
* ... rooms in LSB *or* HH?
* substitution:
```{r regsub}
import re
x = "(Smith, John), (Schmo, Joe)"
re_pat = "\(([A-Z][a-z]+), ([A-Z][a-z.]*)\)"
print(re.sub(re_pat,"(\\2 \\1)",x))
```
* in general if you want something to be interpreted **literally** in a regular expression (e.g. you want to look for `+` or `(` and not have them interpreted in their regular-expression sense), you need to use a *backslash* `\` as an "escape" character: for example, `"a\+b"` will look for the letter `a`, followed by a plus sign, followed by the letter `b` -- *not* one or more `a`s, followed by a `b` ([xkcd](http://xkcd.com/234/))
