---
title: "Benford's Law; file I/O in Python"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine="python",engine.path='python3')
```

## Benford's law

*Benford's law* describes the (surprising) distribution of first digits of many different sets of numbers.
Read it about it on [Wikipedia](http://en.wikipedia.org/wiki/Benford%27s_law) or
[MathWorld](http://mathworld.wolfram.com/BenfordsLaw.html)

We'll write a Python function `benford_count` that tabulates the occurrence of digits from a set of numbers

- we'll write the following functions:
    - `init_dict()`: initialize a dictionary indexed by digits. (Overkill for this example.)
	    - `for` loop
		- or `zip()` function: in general, this "flips" a set of tuples, combining the first elements, second elements, ... The results come back as a "zip object", but can be converted via `list()` or `tuple()`, e.g.
```{r zip1}
zz = zip(("a","b","c"),(1,2,3),("A","B","C"))
print(tuple(zz))
```

- <!-- -->
    - `update_dict(dd,n)`: account first digit of number `n` in the dictionary
	- `benford_loop(d)`: loop over a set of digits `d`
    - test!
- But where do we get our numbers from?

## file I/O

### A bunch of ways to read a file

[Reference](https://docs.python.org/3/tutorial/inputoutput.html)

- `f = open(filename,mode)` (mode = `r` for reading, `w` for writing, `a` for appending) - text by default; returns a **file object**
```{r openfile}
f = open("test.txt")
print(f)
```
- `f.read(size)`; reads `size` characters (`f.read()` reads the whole file), unless we are at the end of the file; then returns `''`
```{r readfile}
f = open("test.txt")
print(f.read())
```
- includes **new lines** (`\n`)
- or we can read a line at a time, or read lines into a list:
```{r readline}
f = open("test.txt")
L = f.readline() ## read one line
print(L)
print(L,end="")  ## print normally *appends* a newline
print(repr(L))   ## repr() prints a *representation* of the line
L2 = f.readlines()
print(L2)
```
- you can use a `for` loop: `for L in f: do_something_with(L)`.
- the file object keeps track of how much has been read

### `next()`, and more flow control

- what if we want to a little more control? **next()** function, works on any **iterable** (tuples, lists, ranges, files ...)
```{r loop}
f = open("test.txt")
L = next(f) ## read one line
print(L)
```
- when you get to the end of a file (or a list or whatever) and try to use `next()` you get a `StopIteration` error; use `try`/`except` to handle it safely
```{r nextloop}
f = open("test.txt")
finished = False
while not finished:
   try:
      L = next(f)
   except StopIteration:
      finished = True
```
- `try` is a *general* way to handle errors safely
- While we're catching up: `break` is also useful. It bails out prematurely of any `while` or `for` loop, e.g.

```{r break1}
for i in range(5):
   print(i)
   if i==3:
      break
```
- a standard idiom for doing something until it works:
```{r break2,eval=FALSE}
while True:
   try:
      x = int(input("enter a number: "))
      break
   except ValueError:
      print("Try again!")
```

### More I/O details

- getting rid of pesky newlines: `.strip()` method for strings (gets rid of leading and trailing *whitespace*)
```{r strip}
f = open("test.txt")
L = f.readline() ## read one line
print(repr(L))
print(repr(L.strip()))
```
- breaking lines into words: `.split()` method for strings
```{r split}
f = open("test.txt")
L = f.readline() ## read one line
LL = L.strip().split(" ")
print(LL)
```

### zip()

(optional!)

- `zip(a,b,c)`, where `a`, `b`, `c` are lists or tuples of the same length, puts together the first elements, second elements, third elements ... for example
```{r zip}
print(zip(("A","B","C"),("a","b","c"),(1,2,3)))
```
- this comes in handy sometimes.  It makes a handy way to invert a dictionary: `dict(zip(d.values(),d.keys()))`

## More i/o stuff

- `import os` in order to find out working directory (`os.getcwd()`), or set the directory `os.chdir(newpath)`; use full path or use (e.g.) `..` to go up one level
- opening a URL from the web `import urllib.request as ur`; `ur.urlopen(url)`
- to read a text file: `io.TextIOWrapper()`
- to read a CSV file: `import csv`, use `csv.reader()`

## Regular expressions

- `import re`
- simple version: look for or replace a string
    - `bool(re.search(pattern,text))` (in this case same as `pattern in text`)
	- `re.sub(text,pattern,replacement)`: e.g. `re.sub("hello","ll","")`
- **patterns**

----------------------------------------------------------
   Pattern   Description                                           
-----------  ---------------------------------------------
 `^`         Beginning of line
 
 `$`	     End of line    
 
 `.`         Any single character except newline
 
 `[...]`	 Any single character in brackets         
 
 `[^...]`	 Any single character **not** in brackets 
 
 `re*`	     0 or more occurrences of preceding expression
 
 `re+`	     1 or more occurrence of preceding expression
 
 `re?`	     0 or 1 occurrence of preceding expression  
 
 `re1|re2`   match `re1` or `re2`
 
 `()`        grouping
 
----------------------------------------------------------

- e.g. `[0-9]+` is a number; `[0-9.]+` includes decimal point; `[0-9.]+, [0-9.]+` is two words separated by a comma and a space
- e.g. remove commas from a string via `re.sub(text,",","")`
- remove punctuation via `re.sub(text,"[,;.!?]","")` *or* `re.sub(text,"["+string.punctuation+"]","")`


<!-- 
   http://testingbenfordslaw.com/
   scrape:
   http://www.atlasoftheuniverse.com/stars.html
-->
