# Introduction

## Administrative trivia

* Instructor: Ben Bolker
    * `bolker@mcmaster.ca`: please include `1mp3` in Subject:
	* `http://www.math.mcmaster.ca/bolker`
	* HH 314 (sometimes LSB 336); office hours TBA
* TA: Jake Szamosi 
* Grading:
    * midterm 20%
	* final (take-home?)    30%
	* weekly assignments 30%
	* project  20%
* Laptop policy
* Course material on [Github](https://github.com/bbolker/math1mp) and Avenue
* Expectations of professor and students
* Textbook (none); see [resources](https://github.com/bbolker/math1mp/misc/resources.md)

# More interesting stuff

## Using computers in math and science

* math users vs. understanders vs. developers
* develop conjectures; draw pictures; write manuscripts
* mathematical proof (e.g. [four-colo(u)r theorem](http://en.wikipedia.org/wiki/Four_color_theorem) and [other examples](http://math.stackexchange.com/questions/1084230/what-are-some-theorems-that-currently-only-have-computer-assisted-proofs)); computer algebra
* applied math: cryptography, tomography, logistics, finance, ...
* applied statistics: bioinformatics, Big Data/analytics, ...
* discrete vs. continuous

## Fun!

### Hello, world

```{r opts,message=FALSE,echo=FALSE}
library("knitr")
```

```{r hello_world,engine='python'}
print('hello, python world!')
```

Python as a fancy calculator:
```{r calc,engine='python'}
print(62**2*27/5+3)
```

## Interlude: about Python

* [scripting](http://en.wikipedia.org/wiki/Scripting_language); high-level; glue; general-purpose; flexible
* relatively modern (1990s; Python 3, 2008)
* currently the [8th most popular computer language](http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html) overall; [most popular for teaching](http://cacm.acm.org/blogs/blog-cacm/176450-python-is-now-the-most-popular-introductory-teaching-language-at-top-us-universities/fulltext)
* well suited to mathematical and scientific programming ([NumPy](http://www.numpy.org); [SciPy](http://www.scipy.org))
* ex.: [Sage](http://www.sagemath.org); [BioPython](http://www.biopython.org)

### the Julia set

Suppose we iterate $z_{n+1}=z_{n}^2+c$ for some complex number $c$, for values of $z$ in the complex plane.
The [Julia set](https://en.wikipedia.org/wiki/Julia_set) is the boundary of the set for which the iterations do *not* go off to infinity. (*What is the Julia set for $c=0+0i$?*)

The basic method for generating pretty pictures is:

* pick a value of $c$
* for lots of different values of $z_0$
    * keep calculating $z_{n+1}=z_n^2+c$ until $\textrm{mod}(z_{n+1})$ is greater than some threshold $T$
	* record $T$
* translate values of $T$ into some colour scale and plot the results	

Complex arithmetic is built into Python (*What is $(2+3i)^2$=*`(complex(2,3))**2`?)

Iterate by hand ...
```{r julia1,engine='python'}
print(complex(0.5,0.5)**2+complex(0,0.65))
print((complex(0.5,0.5)**2+complex(0,0.65))**2+complex(0,0.65))
print(((complex(0.5,0.5)**2+complex(0,0.65))**2+complex(0,0.65))**2)
```

Use **assignments** to simplify ...
```{r julia2,engine='python'}
c=complex(0,0.65)
z0=complex(0.5,0.5)
T=10
z1=z0**2+c
z2=z1**2+c
z3=z2**2+c
print(abs(z3)<T)
```

[Julia set program](../code/julia.py) (modified from [Ted Burke](https://batchloaf.wordpress.com/2013/02/10/creating-julia-set-images-in-python/))


## Interfaces

* command line
* [programming editor]()
* IDE
* GUI

## Assignment

* variable name rules
* variable name conventions (camelCase etc.)


