---
title: "1MP3 Midterm Review 2016"
date: "Tuesday February 9, 2016"
author: Dexter barrows
output:
  html_document:
    theme: cosmo
    highlight: tango
---

```{r, echo = FALSE, warning = FALSE}
library("knitr")
opts_chunk$set(tidy = FALSE, engine = 'python',engine.path = 'python3')
```

[Python documentation reference](https://docs.python.org/3/tutorial).

***


## Strings

Arrays of characters, can be used to hold single characters, words, sentences, user input/output, etc. Strings are declared using single or double quotes.

```{r}
c   = "a"
wd  = "moxie"
sen = "Your midterm is on Friday!"
print(c)
print(wd)
print(sen)
```

They can be easily concatenated or repeated.

```{r}
print("hi" + "there")
print(3*"blah")
```

#### Building Strings

You can 'build' strings by creating an empty string and adding to it

```{r}
s = ""
s += "What's"
s += " "
s += "up?"
print(s)
```

***

## Lists

Lists are used to store arrays of any type. They can all be the same type (for example `int`), or differing types. They can also be added together like strings.

```{r}
numlist  = [1, 2, 3, 4, 5, 525600]
wordlist = ['this', 'is', 'a', 'list', 'of', 'words']
print(numlist)
print(wordlist)
print(numlist + wordlist)
```

#### Nested Lists

Lists can be nested too (lists in lists).

```{r}
nestlist = [[1,2,3],[4,5,6],[]]
print(nestlist)
```

Useful methods for working with lists include `append()` for adding a **single** element to the end of a list, *no matter what type it is*.

```{r}
numlist = [1, 2, 3]
print(numlist)
numlist.append(4)
print(numlist)
numlist.append([5])
print(numlist)
```

An alternative method, `extend()`, can be used to concatenate lists similar to using the plus (`+`) operator. Note that `extend()` expects its argument to be a list.

```{r}
numlist = [1, 2, 3]
print(numlist)
numlist.extend([4])
print(numlist)
numlist.extend([5, 6, 7])
print(numlist)
```

#### Note on Mutability

If you assign a list variable to another variable, both variables will point **at the same list**.

```{r}
L1 = [1, 2, 3, 4, 5]
L2 = L1
L2.append(55)
print(L1)
print(L2)
```

#### Useful methods for Lists

Some other helpful methods for working with lists (from class notes):

* `x.remove(value)`: removes *first* value
* `x.reverse()` (or `x[::-1]`)
* `x.index(value)`: first occurrence of `value`
* `len(x)`: length

***

## Tuples

Similar to lists, but are **immutable** (more on that later).

```{r}
tup = (1, 2, 3, (4, 5))
print(tup)
```

***

## Indexing and Slicing

Strings, lists, and tuples all support indexing and slicing. This means we can access individual elements or ranges of elements in any of these types. Recall that element **indices start at 0**.

```{r}
nums = [1,2,3,4,5]
s = "Pythons are snakes"
print(nums[0])		# indexing
print(nums[1:4])	# slicing
print(s[-1]) 		# indexing from end
print(s[-3:])		# slicing to end
```

With nested lists/tuples, double indexing is used.

```{r}
tup = ((1,2,3),(4,5))
print(tup[0])
print(tup[0][1])
print(tup[-1][-1])
```

With lists you can assign *to* elements using indexing/slicing, this **does not** work with strings/tuples as they are **immutable**.

```{r}
numlist = [[1, 2, 3], [], [4, 5]]
print(numlist)
numlist[0][2] = 42
print(numlist)
```

***

## If statements

If statements can be used to conditionally execute code. We usually check conditions using logical operators such as

* `==` equal to
* `!=` not equal to
* `<` less than
* `<=` less than or equal to
* `>` greater than
* `>=` greater than or equal to

For example:

```{r}
x = 5
if x < 17:
	print("x is smaller")
```

Alternative conditions can be checked at the same time using `elif` statements, and a default action can be set using `else`.

```{r}
x = 42
if x < 17:
	print("x is smaller")
elif x > 17:
	print("x is bigger")
else:
	print("x is the same")
```

***

## Loops

We can repeatedly carry out sequences of instructions using `for` or `while` loops. Traditionally, `for` loops are used when we know how long the loop needs to run for, and `while` loops are used when we do not ro would rather have the computer figure it out.

```{r}
for i in range(10):
	print(i)
```

```{r}
i = 0
while i**2 < 87:
	print(i)
	i += 2
```


## Functions

We can define functions to store sections of code we want to run repeatedly. The input/output is specified using arguments and return statements

```{r}
def add1(x):
	x += 1
	return(x)

x = 3
print(add1(x))
print(x)
```

Note that `x` is not modified in the above example because it is **immutable**. Immutable types are numbers (`int`/`float`), strings, and tuples. Imporatant: **lists are mutable**.

```{r}
def append4(list):
	list.append(4)
	return(list)

numlist = [1,2,3]
print(append4(numlist))
print(numlist)
```
