---
title: "basics of numeric representations"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
library("runr")
py <- proc_python()
py$start()
knit_engines$set(python = function(options) {
    knitr:::wrap(py$exec(options$code), options)
})
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

## Integers

- stored as binary digits
- `n` digits gives room to store *signed* values between min of $-2^(n-1)$ and  max of $2^(n-1)-1$
- base Python automatically switches from 64-bit to arbitrary-length integers as necessary
- in `numpy` can specify precision of integers, or *unsigned* integers (e.g. `uint8`)
- you should rarely do this
- and [be careful if you do](https://xkcd.com/571/)

[information about types and ranges](http://docs.scipy.org/doc/numpy-1.10.1/user/basics.types.html)

## Floats

- default 64-bit floats: 1 sign bit, 11 bits for exponent (`x`), 52 bits for mantissa (`m`)
- can store up to *approximately* $2^(2^(x-1))$; numbers less than $10^{-324}$ *underflow* to zero; numbers greater than $10^{308}$ give `OverflowError`
- only 52 bits of precision in mantissa; for $x \le -16$, `1+x` *underflows* to `1.0`.
- similar issues occur as long as addends are far apart, e.g. `10**9+10**-8`
- what can you do?
     - more *stable* algorithm (e.g. add items in increasing order)
	 - work on the log scale (i.e. add log values rather than multiplying values)
     - extended/arbitrary precision floats: `decimal` module (built in), `mpmath`    - always be careful comparing floating point:
```{r}
import math
a = math.sqrt(2)
print(a*a - 2)
```

Use something like [this](http://stackoverflow.com/questions/4028889/floating-point-equality-in-python):

```{r}
def approx_equal(a, b, tol=1e-8):
     return abs(a - b) < tol
```


Lots more detail [here](https://docs.oracle.com/cd/E19957-01/806-3568/ncg_goldberg.html)
