---
title: "numpy"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
library("runr")
py <- proc_python()
py$start()
knit_engines$set(python = function(options) {
    knitr:::wrap(py$exec(options$code), options)
})
opts_chunk$set(tidy=FALSE,engine='python',engine.path='python3')
```

### Installation

I am hoping that everyone has successfully followed Dexter's instructions for [installing Anaconda+Python](http://bbolker.github.io/math1mp/notes/install_python.html). If you have, then you should be able to just `import numpy as np` to get access to `numpy`.

### arrays

- `np.array()`: construct an array from an existing list/tuple/list of lists/etc.
- creates a homogeneous numeric (floating point (`float64`) or integer (`int64`))
- 1-D: vectors
    - vectorized/elementwise arithmetic operators!
	     - 	arithmetic now works on *elements*, not on the list: `a+1` creates a new array with 1 added to every element; `a+=1` adds 1 to every element
    - create from list or tuple (specify type explicitly; `np.array(x,dtype)` where `type` is `int`, `str`, `float`, `complex` ...)
```{r dtype}
import numpy as np
a1 = np.array([1,2])
print(a1.dtype)
```
```{r dtype2}
a2 = np.array([1,2],type="float")
print(a2.dtype)
```

-
    - slicing and indexing work the same way as lists/tuples/strings
     - **mutable** like lists/dictionaries, so we can set elements (e.g. `a[1]=0`) but we can `.copy()`	 ...
   	- some useful ways to create arrays directly:
	     - `np.arange()`: like `range`, give start, stop, step, (dtype)
		 - `np.zeros()`: all zeros
		 - `np.ones()`: all ones
		 - `np.linspace(start, stop, num=50, endpoint, dtype)`;

	- multi-dim
	     - create by passing a list of lists/tuple of tuples
	     - index via `a[i,j]` (hurrah)
	     - arithmetic is still elementwise; in particular, multiplication is elementwise (*Hadamard product*)
		 - multi-dimensional slicing:
		      - e.g. `a[:,0:2]` takes the first two columns
			  - how would you get odd rows?
			  - flip the order of columns?
		 - dimensions: `a.shape`  (**note** no parentheses! `shape` is an **attribute**, like `a.dtype`
		 - `a.fill()`: fill an array with a (scalar) value
		 - `a.reshape(shape)`:  change dimensions; **row-major order**
         - `a.transpose()`
		 - `a.flatten()`
		 - `np.concatenate((a1,a2,a3),axis=0)`
    - arrays that do not match in the number of dimensions will be **broadcasted** (i.e. by Python to perform mathematical operations.
	- matrices
	     - 2-D arrays
		 - `dot` function
		 - `np.eye()`
		 - `np.diag(v,k=0)`
		 - `np.linalg` submodule
    - vectorized *logical*/comparison operations and indexing
	     - e.g. `a>0` gives array of `bool`
		 - `a[a>0]` selects only positive elements of `a`
		 - `any` and `all`
	- other useful stuff
	     - `.sum()`, `.prod()`, `.max()`, `.min()`, `.argmax()`, `.argmin()`, `.mean()` (note `axis` argument), `sd()`, `rank()` ...

### examples

- (32) take a 2-dimensional array and scale it so the columns have mean 0 and standard deviation 1
- (18) Create a 3x3 matrix with row values ranging from 0 to 2 (various choices)
- (33) sort an array by the nth column

### to do in class

- calculate the mean of the squares of the natural numbers up to 7
- create a 5 x 5 array with row values ranging from 0 to 1 by 0.2
- create a 3 x 7 array containing the values 0 to 20 and a 7 x 3 array containing the values 0 to 20 and matrix-multiply them: result should be
```
array([[ 273,  294,  315],
       [ 714,  784,  854],
       [1155, 1274, 1393]])
```

References:

- [official docs](http://docs.scipy.org/doc/)
- [useful unofficial intro](http://www.engr.ucsb.edu/~shell/che210d/numpy.pdf)
- [numpy examples](http://www.labri.fr/perso/nrougier/teaching/numpy.100/index.html)
