---
title: "Dictionaries and tuples"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine="python",engine.path='python3')
```

[Reference](https://docs.python.org/3/tutorial/datastructures.html);
[reference](http://www.sthurlow.com/python/lesson06/)

## Tuples

* simple; **non-mutable** version of lists
* faster, safer
* can be expressed as  `x, y, z` (or `(x,y,z)`, probably clearer)
* empty tuple: `()`
* tuple with one element: `(x,)`
* can do many of the same things as with lists
```{r tuple_stuff}
x = (1,4,"a",3)
print(x[1])   ## indexing
print(x[2:])  ## slicing
print(x+(3,)) ## appending
print(x[:2] + (3,) + x[2:]) ## insertion in the middle
x.index(4)    ## indexing
"z" in x      ## looking for stuff
x.count(4)    ## count
```
* you *can't* modify the existing tuple at all (deletion, modification)
* unpacking: `x,y,z = t`
* swapping:  `(a,b) = (b,a)`
* useful as the return value of functions; safe, and can be unpacked
* convert to/from lists (`tuple()`, `list()`)
```{r tuple_2}
x = (1,2,3)
def modify(x):
    y = list(x)
    y[0] = "a"
    return(tuple(y))

print(modify(x))
print(x)
```

## Dictionaries

* **unordered**
* efficient: *hashing*
```{r dict1}
d = dict(A=1,B=2,C=3)
print(d["A"])
## d[1] won't work
print("A" in d)
print(1 in d)
print(d.values())
print(d.keys())
x = (("A",1),("B",2))
dict(x)
```
* can add and remove entries from a dictionary
```{r}
d = dict(A=1,B=2,C=3)
d["D"]=5
del d["A"]
d.pop("C")
```
* alternate syntax:
```{r dict2}
{"A":1, "B":2, "C":3}  ## need quotes around keys
{} ## empty dictionary
```
* loop over *keys* in a dictionary:
```{r dict3,eval=FALSE}
d = dict(A=1,B=2,C=3)
for i in d.keys():
    ## do something
```

More Roman numerals!

## Benford's law

*Benford's law* describes the (surprising) distribution of last digits of many different sets of numbers.
Read it about it on [Wikipedia](http://en.wikipedia.org/wiki/Benford%27s_law) or
[MathWorld](http://mathworld.wolfram.com/BenfordsLaw.html)

We'll write a Python function `benford_count` that tabulates the occurrence of digits from a set of numbers

* Set up a dictionary
    * do this with a `for` loop
	* or by passing it a list or tuple of lists/tuples
	* using the `zip()` function
* Function for updating dictionary
* (Maybe) function for removing commas from a string (using `str.replace()`)	

* Where do we get our numbers from?
* How do we iterate through a file?
* How do we open a file from the internet?
* How do we deal with a CSV file?

For example, if the text file contained only three lines
```
1,234
123
12
```
then the correct result would be `[0,0,1,1,1,0,0,0,0,0]` (i.e. the last digits per line in this file comprise one 2, one 3, and one 4; there are zero of every other digit).

* working through a file: `.next()` method


