---
title: "Dictionaries and tuples"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine="python",engine.path='python3')
```

[Reference](https://docs.python.org/3/tutorial/datastructures.html);
[reference](http://www.sthurlow.com/python/lesson06/)

## Tuples

* simple; **non-mutable** version of lists
* faster, safer
* can be expressed as  `x, y, z` (or `(x,y,z)`, probably clearer)
* empty tuple: `()`
* tuple with one element: `(x,)`
* can do many of the same things as with lists
```{r tuple_stuff}
x = (1,4,"a",3)
print(x[1])   ## indexing
print(x[2:])  ## slicing
print(x+(3,)) ## appending
print(x[:2] + (3,) + x[2:]) ## insertion in the middle
x.index(4)    ## indexing
"z" in x      ## looking for stuff
x.count(4)    ## count
```
* you *can't* modify the existing tuple at all (deletion, modification)
* unpacking: `x,y,z = t`
* swapping:  `(a,b) = (b,a)`
* useful as the return value of functions; safe, and can be unpacked
* convert to/from lists (`tuple()`, `list()`)
```{r tuple_2}
x = (1,2,3)
def modify(x):
    y = list(x)
    y[0] = "a"
    return(tuple(y))

print(modify(x))
print(x)
```

## Dictionaries

* **unordered**
* efficient: *hashing*
```{r dict1}
d = dict(A=1,B=2,C=3)
print(d["A"])
## d[1] won't work
print("A" in d)
print(1 in d)
print(d.values())
print(d.keys())
x = (("A",1),("B",2))
dict(x)
```
* can add and remove entries from a dictionary
```{r}
d = dict(A=1,B=2,C=3)
d["D"]=5
del d["A"]  ## or d.pop("A")
```

More Roman numerals!

## Benford's law

*Benford's law* describes the (surprising) distribution of last digits of many different sets of numbers.
Read it about it on [Wikipedia](http://en.wikipedia.org/wiki/Benford%27s_law) or
[MathWorld](http://mathworld.wolfram.com/BenfordsLaw.html)

We'll write a Python function `benford_count` that takes the name of a file (that will contain one number per line), opens the file, and tabulates the occurrences of last digits on each line.

* using `open()` to open a file; we can then iterate through the lines with a `for` function (or a `while`; )
* **NOTE** that if you follow the instructions below to read and process the file, the last character of each line is a *newline* character; if you want the last *digit* you need to get the second-to-last character of the string, or use `str.strip()`.

```{r eval=FALSE}
ff = open("benford.txt")
for line in ff:
    ## code
```

For example, if the text file contained only three lines
```
1,234
123
12
```
then the correct result would be `[0,0,1,1,1,0,0,0,0,0]` (i.e. the last digits per line in this file comprise one 2, one 3, and one 4; there are zero of every other digit).

Keep in mind that (1) what you read out of the file will be a *string* and (2) you can turn a digit into an integer using `int()`, e.g.

```{r}
x = "7"
print(int(x))
```
