---
title: "debugging"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE}
library("knitr")
source("python_engine3.R")
opts_chunk$set(tidy=FALSE,engine='python3')
```

# Debugging

## Errors

### Types of errors

* syntax errors vs. logic errors

* failure modes:
    * obvious failure
        * program stops with an error partway through: [bad mandelbrot #0](../code/mandelbrot/noinit.py)
        * Python crashes
        * machine crashes
		* program never stops (infinite loop)
    * wrong answer
	    * always
        * occasionally

* infinite loops: [bad mandelbrot #1](../code/mandelbrot_infloop.py)
* incorrect operator precedence, e.g. $\Delta \textrm{fahrenheit} = \Delta \textrm{Celsius} \times 1.8$
```{r FCconv,eval=FALSE}
fahrdiff = celsius_high - celsius_low * 1.8
```
* off-by-one error ("fencepost problem")
* incorrect code inside/outside loops: 
   * [bad mandelbrot #2](../code/mandelbrot_badloop1.py)
   * [bad mandelbrot #3](../code/mandelbrot_badloop2.py)
   * [bad mandelbrot #4](../code/mandelbrot_badloop3.py)
* array index error (outside bounds?)

### Error messages

* error messages are *trying* to tell you something
* Google error messages (with quotation marks)  [XKCD]

## Debugging

* *brute-force logic*: stare at your code, try to figure out what's wrong    
(test cases really help: why is it failing in one specific situation?)
* flow charts, *pseudocode*
* `print()`
* interactive tracing
* debugging tools (breakpoints/watchpoints/watches)
* checkpointing

## Searching for/asking for help

### Searching for help

* Google
* be as specific as possible
* browse

### Asking for help

* reproducible/minimal workable examples
    * right amount of context
	* "how to ask" ([StackOverflow](http://stackoverflow.com/help/how-to-ask))
* browse/lurk in forums first!
* tone
* where:
    * forums
    * StackOverflow

**EXAMPLE**

* http://stackoverflow.com/questions/1623039/python-debugging-tips
* https://www.udacity.com/course/cs259
* http://www.cs.yale.edu/homes/aspnes/pinewiki/C%282f%29Debugging.html
* http://www.cs.cf.ac.uk/Dave/PERL/node149.html
* http://space.wccnet.edu/~pmillis/cps120/presentations/program_logic_errs.ppt


