---
title: "More functions etc. examples"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine="python",engine.path='python3')
```

## More examples

### conversion to/from Roman numerals

* Based on [CSE 231 project "Shannon's Throbac"](http://www.cse.msu.edu/~cse231/PracticeOfComputingUsingPython/03_Strings/Shannon/Project04.pdf)
* The rules:
    * (I,V,X,L,C,D,M) $\leftrightarrow$ (1,5,10,50,100,500,1000)
	* *subtraction rule*: I, X, or C preceding the next two elements correspond to subtraction (e.g. IV=4, XC=90)
* [string methods](https://docs.python.org/3/library/stdtypes.html#str) are useful, especially `in`, `s.count()`, `s.replace()` are useful
* so is *integer division* and *modulo* operations, `//` and `%`
* we should write tests first, e.g.
```{r roman1,eval=FALSE}
rom = ["I","II","III","IV","V","VI","VII","VIII","IX",
     "X","XI","XIV","XV","LX","XC","XCXLI","CCC"]
int = [1,2,3,4,5,6,7,8,9,10,11,14,15,60,90,131,300]
for i in range(len(rom)):
    print(rom_to_int(rom[i])==int[i])
```
* we can try to write this as a gigantic `if` statement, or we can try to make the rules more compact
* which direction is easier?
          
### compute sine/cosine

* [CSE project](http://www.cse.msu.edu/~cse231/PracticeOfComputingUsingPython/02_Control/Trig/project04.pdf)
* power series approximation:
$$
\begin{split}
\sin(x) & = x-\frac{x^3}{3!}+\frac{x^5}{5!}-\frac{x^7}{7!} + \dots \\
& = \sum_{k=0} (-1)^k \frac{x^{2k+1}}{(2k+1)!}
	\end{split}
$$
$$
\begin{split}
\cos(x) & = 1-\frac{x^2}{2!}+\frac{x^4}{4!}-\frac{x^6}{6!} + \dots \\
        & = \sum_{k=0} (-1)^k \frac{x^{2k}}{(2k)!}
\end{split}
$$
* stop when the absolute value of the next term is $<10^{-8}$ (`1e-8` in computer notation)
* divide $x$ by $\pi$, take the remainder (modulo: `%`)
* write tests first (what are some easy cases, using `math.pi`?)
* use `math.factorial`? use the factorial function we already wrote? augment terms as we go along?
* after we write the two functions, can we combine them?
* extra credit/puzzle: what do we get if we interpolate between these two functions?

### letter-scrambling

*  [CSE project](http://www.cse.msu.edu/~cse231/PracticeOfComputingUsingPython/04_Functions1/ScrambledWords/project06.pdf)
* scramble words within a text
* start with some simpler cases
    * reverse every word in a line
    * again, [string methods](https://docs.python.org/3/library/stdtypes.html#str) -- especially `str.split()`
	* now read from a file 
* test cases? easier with just a few lines	
* harder: preserve punctuation at the end of the word
* harder: shuffle
   * `from numpy.random import shuffle`
* split words and join them: `list()`, `''.join()`
* [test case](../misc/darwin.txt) 
```{r}
f = open("../misc/darwin.txt")
for line in f:
   print(line)
```
