---
title: "functions (part 2) and modules"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine="python",engine.path='python3')
```

## Functions calling functions

Functions can call functions (even themselves, **recursively**):

```{r}
def factorial(x):
    if (x==1):
        return(1)
    return(x*factorial(x-1))

factorial(5)
```

What happens if we forget to put in the `if` clause?

- More often, functions call functions
- You can pass anything to a function as an argument (even a function!)

```{r}
def repeat_fun(f,x,n):
    for i in range(n):
        x=f(x)
    return(x)

def sqr(x):
    return(x*x)

repeat_fun(sqr,3,3)
```

### Scope

* Where does Python look for things?
* What happens here?
```{r}
z = 1
def add_z(x):
    return(x+z)

add_z(z)
```
* **LEGB** (Local, Enclosing, Global, Built-in)
    * *Local*: symbols defined in the function, and arguments
	* *Enclosing*: symbols defined *in the function within which this function was defined*
	* *Global*: elsewhere in the file/module
	* *Built-in*: Python keywords
	

# Modules

### importing

* `import`
* refer to functions via module prefix
* `import VeryLongModuleName as vlmn`: use abbreviation


### finding out about

* [official modules](https://docs.python.org/3/py-modindex.html)
* [list of useful module](https://wiki.python.org/moin/UsefulModules)
* `math`, `cmath`, `re`, `random`, `numpy`, `scipy`, `matplotlib`, `timeit`
