---
title: "functions (part 2) and modules"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
author: Ben Bolker
---

```{r opts,message=FALSE,echo=FALSE,warning=FALSE}
library("knitr")
opts_chunk$set(tidy=FALSE,engine="python",engine.path='python3')
```

## Functions calling functions

- You can pass anything to a function as an argument (even a function!)

```{r}
def repeat_fun(f,startval,n):
    """Given a function f and a starting value startval,
    apply the function n times (each time using the previous
    result as input
    """
    y = startval
    for i in range(n):
        y=f(y)
    return(y)

def sqr(x):
    return(x*x)

repeat_fun(sqr,3,3)
```

### Scope

- Where does Python look for things?
- What happens here?
```{r}
z = 1
def add_z(x):
    return(x+z)

add_z(z)
```
- **LEGB** (Local, Enclosing, Global, Built-in)
    - *Local*: symbols defined in the function, and arguments
	- *Enclosing*: symbols defined *in the function within which this function was defined*
	- *Global*: elsewhere in the file/module
	- *Built-in*: Python keywords
	
# Modules

Collections of functions you might want to use.

### importing

- `import`
- refer to functions via module prefix
- `import VeryLongModuleName as vlmn`: use abbreviation
- can import just one or two functions: `from math import sqrt, log`
- can import everything (but usually don't): `from <module> import *`
- can import *your own modules* (i.e., functions in a `.py` file)


### finding out about modules

- `help("modulename")`
- [official modules](https://docs.python.org/3/py-modindex.html)
- [list of useful modules](https://wiki.python.org/moin/UsefulModules)
- some modules we will definitely be using:
    - `math`: basic math functions
    - `matplotlib`: drawing pictures
    - `random`: picking random numbers
    - `numpy`: numerical computation in general (e.g. linear alg and calculus)
    - `pandas`: data analysis
- slightly less useful but maybe using:
    - `nose`: code testing framework
    - `scipy`: even more scientific computing tools
    - `cmath`: math functions handling complex numbers
    - `re`: regular expressions
    - `sympy`: symbolic computation
    - `timeit`: how long does my code take?

